# Deriving Point Estimates

In this chapter, I discuss three methods to derive point estimators of quantities of interest: , Bayes' inference, method of moments (briefly), and maximum likelihood.

## Bayesian Inference

Bayesian inference follows a simple recipe:

1. Choose a distribution for the data.
1. Choose a distribution to describe your prior beliefs.
1. Update the prior distribution upon observing the data by computing the posterior distribution. 

### Mechanics

Suppose a random sample from a distribution $f(x; \theta)$ that depends on the unknown parameter $\theta$. 

Bayesian inference models our *beliefs* about the unknown parameter $\theta$ as a distribution and answers the question: what should we believe about $\theta$, given the observed samples $x = \{x_1, x_2, ..., x_n\}$ from $f(x; \theta)$. These beliefs are simply the conditional distribution $f(\theta \mid x)$.

By Bayes' rule, $\displaystyle f(\theta \mid x) = \frac{f(x \mid \theta)f(\theta)}{f(x)} = \frac{f(x \mid \theta)f(\theta)}{\displaystyle \int_{-\infty}^\infty f(x \mid \theta)f(\theta) d\theta}$.

$$
\displaystyle \underbrace{f(\theta \mid x)}_{\text{posterior}} = \frac{\overbrace{f(x \mid \theta)}^{\text{likelihood}} \times \overbrace{f(\theta)}^{\text{prior}}}{\displaystyle \underbrace{\int_{-\infty}^\infty f(x \mid \theta)f(\theta) d\theta}_{\text{normalizing constant}}}
$$
There are four parts to a Bayesian analysis.

1. $f(\theta \mid x)$. "The posterior;" what we're trying to find. This distribution models our beliefs about parameter $\theta$ given the data $x$. 
1. $f(x \mid \theta)$. "The likelihood." This distribution model conditional density/probability of the data $x$ given the parameter $\theta$. We need to invert the conditioning in order to find the posterior.
1. $f(\theta)$. "The prior;" our beliefs about $\theta$ prior to observing the sample $x$. 
1. $f(x) =\int_{-\infty}^\infty f(x \mid \theta)f(\theta) d\theta$. A normalizing constant. Recall that the role of the normalizing constant is to force the distribution to integrate or sum to one. Therefore, we can safely ignore this constant until the end, and then find proper normalizing constant. 

It's convenient to choose a **conjugate** prior distribution that, when combined with the likelihood, produces a posterior from the same family. 

As a running example, we use the **toothpaste cap problem**:

> We have a toothpaste cap--one with a wide bottom and a narrow top. We're going to toss the toothpaste cap. It can either end up lying on its side, its (wide) bottom, or its (narrow) top. 

> We want to estimate the probability of the toothpaste cap landing on its top.

> We can model each toss as a Bernoulli trial, thinking of each toss as a random variable $X$ where $X \sim \text{Bernoulli}(\pi)$. If the cap lands on its top, we think of the outcome as 1. If not, as 0. 

> Suppose we toss the cap $N$ times and observe $k$ tops. What is the posterior distribution of $\pi$?

#### The Likelihood {#likelihood}

According to the model $f(x_i \mid \pi) = \pi^{x_i} (1 - \pi)^{(1 - x_i)}$. Because the samples are iid, we can find the *joint* distribution $f(x) = f(x_1) \times ... \times f(x_N) = \prod_{i = 1}^N f(x_i)$. We're just multiplying $k$ $\pi$s (i.e., each of the $k$ ones has probability $\pi$) and $(N - k)$ $(1 - \pi)$s (i.e., each of the $N - k$ zeros has probability $1 - \pi$), so that the $f(x | \pi) = \pi^{k} (1 - \pi)^{(N - k)}$.

$$
\text{the likelihood:  } f(x | \pi) = \pi^{k} (1 - \pi)^{(N - k)}, \text{where } k = \sum_{n = 1}^N x_n \\
$$

#### The Prior

The prior describes your beliefs about $\pi$ *before* observing the data.

Here are some questions that we might ask ourselves the following questions:

1. What's the most likely value of $\pi$? *About 0.15.*
1. Are our beliefs best summarizes by a distribution that's skewed to the left or right? *To the right.*
1. $\pi$ is about _____, give or take _____ or so. *Perhaps 0.17 and 0.10.*
1. There's a 25% chance that $\pi$ is less than ____. *Perhaps 0.05.*
1. There's a 25% chance that $\pi$ is greater than ____. *Perhaps 0.20*.

Given these answers, we can sketch the pdf of the prior distribution for $\pi$.

```{r echo=FALSE, fig.height=3, fig.width=6, message=FALSE, warning=FALSE}
library(tidyverse)
#library(xkcd)
library(extrafont)
loadfonts()

set.seed(1234)
a <- 3

b <- 15
s <- rbeta(200, a, b)
gg_data <- tibble(s)
ggplot(gg_data, aes(x = s)) + 
  geom_density() + 
  theme_minimal() + 
  scale_x_continuous(limits = c(0, 1)) +
  annotate(geom = "text", x = .35, y = 1, 
           label = "probably not larger than 0.3", 
           family = "Humor Sans", 
           hjust = "left", 
           color = "grey50") + 
  annotate(geom = "segment", x = .34, xend = 0.30, y = 0.9, yend = 0.5, 
           color = "grey50") + 
  annotate(geom = "text", x = .05, y = 1.5, 
           label = "might be close to zero", 
           family = "Humor Sans", 
           hjust = "left", 
           color = "grey50") + 
  annotate(geom = "segment", x = .1, xend = 0.03, y = 1.3, yend = 0.3, 
           color = "grey50") + 
    annotate(geom = "text", x = 0.18, y = 5.1, 
           label = "most likely around 0.12", 
           family = "Humor Sans", 
           hjust = "left", 
           color = "grey50") + 
  annotate(geom = "segment", x = .14, xend = 0.17, y = 4.9, yend = 5.05, 
           color = "grey50") + 
  theme(text = element_text(size = 12, family = "Humor Sans"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) + 
  scale_y_continuous(breaks = NULL) + 
  labs(title = "A Sketch of Our Prior Beleifs About pi",
       x = "pi",
       y = "Prior Density") 

```

Now we need to find a density function that matches these prior beliefs. For this Bernoulli model, the *beta distribution* is the conjugate prior. While a conjugate prior is not crucial in general, it makes the math much more tractable. 

So then what beta distribution captures our prior beliefs?

There's a code snippet [here](https://gist.github.com/carlislerainey/45414e0d9f22e4e1960449402e6a8048) to help you explore different beta distributions.

After some exploration, we find that setting the parameters $\alpha$ and $\beta$ of the beta distribution to 3 and 15, respectively, captures our prior beliefs about the probability of getting a top.

```{r}
ggplot(data_frame(x = seq(0, 1, length.out = 100))) + 
  stat_function(fun = dbeta, args = list(shape1 = 3, shape2 = 15)) + 
  labs(title = "pdf of the beta(3, 15) distribution", x = "pi", y = "prior density")
```
The pdf of the beta distribution is $f(x) = \frac{1}{B(\alpha, \beta)} x^{\alpha - 1}(1 - x)^{\beta - 1}$. Remember that $B()$ is the beta function, so $\frac{1}{B(\alpha, \beta)}$ is a constant.

Let's denote our chosen values of $\alpha = 3$ and $\beta = 15$ as $\alpha^*$ and $\beta^*$. As we see in a moment, it's convenient distinguish the parameters in the prior distribution from other parameters.

$$
\text{the prior:  }  f(\pi) = \frac{1}{B(\alpha^*, \beta^*)} \pi^{\alpha^* - 1}(1 - \pi)^{\beta^* - 1}
$$

#### The Posterior

Now we need to compute the posterior by multiplying the likelihood times the prior and then finding the normalizing constant.
$$
\text{the posterior: } \displaystyle \underbrace{f(\pi \mid x)}_{\text{posterior}} = \frac{\overbrace{f(x \mid \pi)}^{\text{likelihood}} \times \overbrace{f(\pi)}^{\text{prior}}}{\displaystyle \underbrace{\int_{-\infty}^\infty f(x \mid \pi)f(\pi) d\pi}_{\text{normalizing constant}}} \\
$$
Now we plug in the likelihood, plug in the prior, and denote the normalizing constant as $C_1$ to remind ourselves that it's just a constant.

$$
\text{the posterior: } \displaystyle \underbrace{f(\pi \mid x)}_{\text{posterior}} = \frac{\overbrace{\left[ \pi^{k} (1 - \pi)^{(N - k) }\right] }^{\text{likelihood}} \times \overbrace{ \left[ \frac{1}{B(\alpha^*, \beta^*)} \pi^{\alpha^* - 1}(1 - \pi)^{\beta^* - 1} \right] }^{\text{prior}}}{\displaystyle \underbrace{C_1}_{\text{normalizing constant}}} \\
$$

Now we need to simplify the right-hand side. 

First, notice that the term $\frac{1}{B(\alpha^*, \beta^*)}$ in the numerator is just a constant. We can incorporate that constant term with $C_1$ by multiplying top and bottom by $B(\alpha^*, \beta^*)$ and letting $C_2 = C_1 \times B(\alpha^*, \beta^*)$.

$$
\text{the posterior: } \displaystyle \underbrace{f(\pi \mid x)}_{\text{posterior}} = \frac{\overbrace{\left[ \pi^{k} (1 - \pi)^{(N - k) }\right] }^{\text{likelihood}} \times  \left[ \pi^{\alpha^* - 1}(1 - \pi)^{\beta^* - 1} \right] }{\displaystyle \underbrace{C_2}_{\text{new normalizing constant}}} \\
$$

Now we can collect the exponents with base $\pi$ and the exponents with base $(1 - \pi)$.

$$
\text{the posterior: } \displaystyle \underbrace{f(\pi \mid x)}_{\text{posterior}} = \frac{\left[ \pi^{k} \times \pi^{\alpha^* - 1} \right] \times  \left[ (1 - \pi)^{(N - k) } \times (1 - \pi)^{\beta^* - 1} \right] }{ C_2} \\
$$
Recalling that $x^a \times x^b = x^{a + b}$, we combine the powers.

$$
\text{the posterior: } \displaystyle \underbrace{f(\pi \mid x)}_{\text{posterior}} = \frac{\left[ \pi^{(\alpha^* + k) - 1} \right] \times  \left[ (1 - \pi)^{[\beta^* + (N - k)] - 1} \right] }{ C_2} \\
$$

Because we're clever, we notice that this is *almost* a beta distribution with $\alpha = (\alpha^* + k)$ and $\beta = [\beta^* + (N - k)]$. If $C_2 = B(\alpha^* + k, \beta^* + (N - k))$, then the posterior is *exactly* a $\text{beta}(\alpha^* + k, \beta^* + [N - k]))$ distribution. 

This is completely expected. We chose a beta distribution for the prior because it would give us a beta posterior distribution. For simplicity, we can denote the parameter for the beta posterior as $\alpha^\prime$ and $\beta^\prime$, so that $\alpha^\prime = \alpha^* + k$ and $\beta^\prime = \beta^* + [N - k]$

$$
\begin{aligned}
\text{the posterior: } \displaystyle \underbrace{f(\pi \mid x)}_{\text{posterior}} &= \frac{ \pi^{\overbrace{(\alpha^* + k)}^{\alpha^\prime} - 1}  \times  (1 - \pi)^{\overbrace{[\beta^* + (N - k)]}^{\beta^\prime} - 1}  }{ B(\alpha^* + k, \beta^* + [N - k])} \\
&= \frac{ \pi^{\alpha^\prime - 1}  \times  (1 - \pi)^{\beta^\prime - 1}  }{ B(\alpha^\prime, \beta^\prime)}, \text{where } \alpha^\prime = \alpha^* + k \text{ and } \beta^\prime = \beta^* + [N - k]
\end{aligned}
$$

This is an elegant, simple solution. To obtain the parameters for the beta posterior distribution, we just add the number of tops (Bernoulli successes) to the prior value for $\alpha$ and the number of not-tops (sides and bottoms; Bernoulli failures) to the prior value for $\beta$.

Suppose that I tossed the toothpaste cap 150 times and got 8 tops.

```{r fig.height=3, fig.width=8, message=FALSE, warning=FALSE}

# prior parameters
alpha_prior <- 3
beta_prior <- 15

# data 
k <- 8
N <- 150

# posterior parameters
alpha_posterior <- alpha_prior + k
beta_posterior <- beta_prior + N - k

# plot prior and posterior
gg_prior <- ggplot() + 
  stat_function(fun = dbeta, args = list(shape1 = alpha_prior, shape2 = beta_prior)) + 
  labs(title = "prior distribution", x = "pi", y = "prior density")
gg_posterior <- ggplot() + 
  stat_function(fun = dbeta, args = list(shape1 = alpha_posterior, shape2 = beta_posterior)) + 
  labs(title = "posterior distribution", x = "pi", y = "posterior density")

library(patchwork)
gg_prior + gg_posterior

```

#### Bayesian Point Estimates

In this section, we want *point estimates*---a best guess at the parameter---not a full posterior distribution. 

We have three options:

1. *The posterior mean*. The posterior mean minimizes a squared-error loss function. That is, the cost of guessing $a$ when the truth is $\alpha$ is $(a - \alpha)^2$. In the case of the beta posterior, it's just $\dfrac{\alpha^\prime}{\alpha^\prime + \beta^\prime}$. For our prior and data, we have $\dfrac{3 + 8}{(3 + k) + (15 + 150 - 8)} \approx 0.065$.
1. *The posterior median*: The posterior median minimizes an absolute loss function where the cost of guessing $a$ when the truth is $\alpha$ is $|a - \alpha|$. Intuitively, there's a 50% chance that $\pi$ falls above and below the posterior median. In the case of the beta posterior, it's just $\dfrac{\alpha^\prime - \frac{1}{3}}{\alpha^\prime + \beta^\prime - \frac{2}{3}}$ (for $\alpha^\prime, \beta^\prime > 1$). For our prior and data, we have $\dfrac{3 + 8 -\frac{1}{3}}{(3 + k) + (15 + 150 - 8) - \frac{2}{3}} \approx 0.064$.
1. *The posterior mode*: The posterior mode is the most likely value of $\pi$, so it minimizes a loss function that penalizes all misses equally. In the case of the beta posterior, it's just $\dfrac{\alpha^\prime - 1}{\alpha^\prime + \beta^\prime - 2}$ (for $\alpha^\prime, \beta^\prime > 1$). For our prior and data, we have $\dfrac{3 + 8 - 1}{(3 + k) + (15 + 150 - 8) - 2} \approx 0.060$.

### Example: Poisson Distribution

Suppose we collect $N$ random samples $x = \{x_1, x_2, ..., x_N\}$ and model each draw as a random variable $X \sim \text{Poisson}(\lambda)$. Find the posterior distribution of $\lambda$ for the gamma prior distribution. Hint: the gamma distribution is the conjugate prior for the Poisson likelihood.

$$
\begin{aligned}
\text{Poisson likelihood: } f(x \mid \lambda) &= \prod_{n = 1}^N \frac{\lambda^{x_n} e^{-\lambda}}{x_n!} \\
&= \displaystyle \left[ \frac{1}{\prod_{n = 1}^N x_n !} \right]e^{-N\lambda}\lambda^{\sum_{n = 1}^N x_n}
\end{aligned}
$$

$$
\text{Gamma prior: } f( \lambda; \alpha^*, \beta^*) = \frac{{\beta^*}^{\alpha^*}}{\Gamma(\alpha^*)} \lambda^{\alpha^* - 1} e^{-\beta^*\lambda}
$$
To find the posterior, we multiply the likelihood times the prior and normalize. Because the gamma prior distribution is the conjugate prior for the Poisson likelihood, we know that the posterior will be a gamma distribution.

$$
\begin{aligned}
\text{Gamma posterior: } f( \lambda  \mid x) &= \frac{\left( \displaystyle \left[ \frac{1}{\prod_{n = 1}^N x_n !} \right]e^{-N\lambda}\lambda^{\sum_{n = 1}^N x_n}\right) \times \left( \left[ \frac{{\beta^*}^{\alpha^*}}{\Gamma(\alpha^*)} \right] \lambda^{\alpha^* - 1} e^{-\beta^*\lambda}\right)}{C_1} \\
\end{aligned}
$$
Because $x$, $\alpha_*$, and $\beta$ are fixed, the terms in square brackets are constant, so we can safely consider those part of the normalizing constant.

$$
\begin{aligned}
&= \frac{\left( \displaystyle  e^{-N\lambda}\lambda^{\sum_{n = 1}^N x_n}\right) \times \left( \lambda^{\alpha^* - 1} e^{-\beta^*\lambda}\right)}{C_2} \\
\end{aligned}
$$
Now we can collect the exponents with the same base.

$$
\begin{aligned}
&= \frac{\left( \lambda^{\alpha^* - 1} \times \lambda^{\sum_{n = 1}^N x_n}\right) \times \left( \displaystyle  e^{-N\lambda} \times e^{-\beta^*\lambda} \right)}{C_2} \\
&= \frac{\lambda^{ \overbrace{\left[ \alpha^* + \sum_{n = 1}^N x_n \right]}^{\alpha^\prime} - 1}  e^{-\overbrace{[\beta^* + N]}^{\beta^\prime}\lambda} }{C_2} \\
\end{aligned}
$$

We recognize this as *almost* a Gamma distribution with parameters $\alpha^\prime = \alpha^* +  \sum_{n = 1}^N x_n$ and $\beta^\prime = \beta^* + N$. Indeed, if $\frac{1}{C_2} = \frac{{\beta^\prime}^{\alpha^\prime}}{\Gamma(\alpha^{\prime})}$, then we have exactly a gamma distribution. 

$$
\begin{aligned}
&= \frac{{\beta^\prime}^{\alpha^\prime}}{\Gamma(\alpha^{\prime})} \lambda^{ \alpha^\prime - 1}  e^{-\beta^\prime\lambda}, \text{where } \alpha^\prime = \alpha^* +  \sum_{n = 1}^N x_n \text{ and } \beta^\prime = \beta^* + N
\end{aligned}
$$

Like the Bernoulli likelihood with the beta prior, the Poisson likelihood withe the gamma prior gives a nice result. We start with values parameters of the gamma distribution $\alpha = \alpha^*$ and $\beta + \beta^*$ so that the gamma prior distribution describes our prior beliefs about the parameters $\lambda$ of the Poisson distribution. Then we add the sum of the data $x$ to $\alpha^*$ and the number of samples $N$ to $\beta^*$ to obtain the parameters of the gamma posterior distribution.

The code below shows the posterior distribution 

```{r fig.height=3, fig.width=8, message=FALSE, warning=FALSE}
# set see to make reproducible
set.seed(1234)

# prior parameters
alpha_prior <- 3
beta_prior <- 3

# create an "unknown" value of lambda to estimate
lambda <- 2

# generate a data set
N <- 5  # number of samples
x <- rpois(N, lambda = lambda)
print(x)  # print the data set

# posterior parameters
alpha_posterior <- alpha_prior + sum(x)
beta_posterior <- beta_prior + N

# plot prior and posterior
gg_prior <- ggplot() + xlim(0, 5) + 
  stat_function(fun = dgamma, args = list(shape = alpha_prior, rate = beta_prior)) + 
  labs(title = "prior distribution", x = "lambda", y = "prior density")
gg_posterior <- ggplot() + xlim(0, 5) + 
  stat_function(fun = dgamma, args = list(shape = alpha_posterior, rate = beta_posterior)) + 
  labs(title = "posterior distribution", x = "lambda", y = "posterior density")
gg_prior + gg_posterior  # uses patchwork package

# posterior mean: alpha/beta
alpha_posterior/beta_posterior

# posterior median: no closed form, so simulate
post_sims <- rgamma(10000, alpha_posterior, beta_posterior)
median(post_sims)

# posterior mode: (alpha - 1)/beta for alpha > 1
(alpha_posterior - 1)/beta_posterior


```

### Remarks

Bayesian inference presents two difficulties. 

1. Choosing a prior. 
   a. It can be hard to actually construct a prior distribution. It's challenging when dealing with a single parameter. It becomes much more difficult when dealing with several or many parameters.
   a. Priors are subjective, so that one researcher's prior might not work for another. 
1. Computing the posterior. Especially for many-parameter problems and non-conjugate priors, computing the posterior can be nearly intractable.

However, there are several practical solutions to these difficulties.

1. Choosing a prior.
    a. We can use a "uninformative" or constant prior. Sometimes, we can use an improper prior that doesn't integrate to one, but places equal prior weight on all values. 
    a. We can use an extremely diffuse prior. For example, if we wanted to estimate the average height in a population in inches, we might use a normal distribution centered at zero with an SD of 10,000. This prior says: "The average height is about zero, give or take 10,000 inches or so."
    a. We can use an informative prior, but conduct careful robustness checks to assess whether the conclusions depend on the particular prior.
    a. We can use a weakly informative prior, that rules places meaningful prior weight on all the plausible values and little prior weight only on the most implausible values. As a guideline, you might create a weakly informative prior by doubling or tripling the SD of the informative prior.
1. Computing the posterior.
    a. While analytically deriving the posterior becomes intractable for most applied problems, it's relatively easy to *sample* from the posterior distribution for many models. 
    a. Algorithms like Gibbs samplers, MCMC, and HMC make this sampling procedure straightforward for a given model.
    a. Software such as Stan make sampling easy to set up and very fast. Post-processing R packages such as tidybayes make it each to work with the posterior simulations.
    
### Exercises 

```{exercise, label="bayes-exponential"}
Suppose we collect $N$ random samples $x = \{x_1, x_2, ..., x_N\}$ and model each draw as a random variable $X \sim \text{expontial}(\lambda)$ with pdf $f(x_n | \lambda) = \lambda e^{-\lambda x_n}$. Find the posterior distribution of $\lambda$ for the gamma prior distribution. Hint: the gamma distribution is the conjugate prior for the exponential likelihood.
```

<details><summary>Hint</summary>
Use the Poisson example from above. Because they share a prior, the math works quite similarly.
</details>


<details><summary>Solution</summary>

$$
\begin{aligned}
\text{exponential likelihood: } f(x \mid \lambda) &= \prod_{n = 1}^N \lambda e^{-\lambda x_n} \\
&= \lambda^N e^{-\lambda \sum_{n = 1}^N x_n}
\end{aligned}
$$
$$
\text{Gamma prior: } f( \lambda; \alpha^*, \beta^*) = \frac{{\beta^*}^{\alpha^*}}{\Gamma(\alpha^*)} \lambda^{\alpha^* - 1} e^{-\beta^*\lambda}
$$
$$
\begin{aligned}
\text{Gamma posterior: } f( \lambda  \mid x) &= \frac{\left(  \lambda^N e^{-\lambda \sum_{n = 1}^N x_n}\right) \times \left( \left[ \frac{{\beta^*}^{\alpha^*}}{\Gamma(\alpha^*)} \right] \lambda^{\alpha^* - 1} e^{-\beta^*\lambda}\right)}{C_1} \\
&= \frac{\left(  \lambda^N e^{-\lambda \sum_{n = 1}^N x_n}\right) \times \left(  \lambda^{\alpha^* - 1} e^{-\beta^*\lambda}\right)}{C_2} \\
&= \frac{ \lambda^{ \overbrace{\left[ \alpha^* + N \right]}^{\alpha^\prime} - 1} e^{- \overbrace{\left[ \beta^* + \sum_{n = 1}^N x_n \right]}^{\beta^\prime} \lambda}}{C_2} \\
& =  \frac{{\beta^\prime}^{\alpha^\prime}}{\Gamma(\alpha^\prime)}  \lambda^{\alpha^\prime - 1} e^{-\beta^\prime\lambda}\text{ where } \alpha^\prime = \alpha^* +  N \text{ and } \beta^\prime = \beta^* + \sum_{n = 1}^N x_n
\end{aligned}
$$

</details>

```{exercise, label="govt-duration-bayes"}
You are a researcher interesting in government duration in parliamentary democracies. You decide to model duration as an exponential distribution. To estimate the parameter $\lambda$ of the exponential distribution, you collect data on the last five UK governments. 

- The first Johnson ministry lasted 142 days. 
- The second May ministry lasted 773 days.
- The first May ministry lasted 333 days.
- The second Cameron ministry lasted 432 days.
- The first Cameron ministry (Cameron-Clegg coalition) lasted 1,823 days.

1. Rescale the data to years--that helps with the interpretation.
1. Create three gamma prior distributions.
    a. One that describes your prior beliefs. 
    a. One that's weakly informative.
    a. One that's as uninformative as possible.
1. Use what you learned from Exercise \@ref(exr:bayes-exponential) to plot each prior and posterior.
1. For each posterior, compute the mean, median, and mode of $\lambda$. Interpret. Hint: $\lambda$ is a rate. If the durations are measured in years, then it's the failures per year. The expected duration in years is $\frac{1}{\lambda}$, but remember that $E \left( \frac{1}{\lambda}\right) \neq \frac{1} {E(\lambda)}$. If you want to find the posterior distribution of the expected duration (rather than the failures/year), then you can simulate many draws of $\lambda$ from the posterior and transform each draw into $\frac{1}{\lambda}$. The mean of the transformed draws is the posterior mean of the expected duration.
1. Assess the model. Do you think we have a good model? Hint: the exponential distribution is memoryless. Does that make sense in this appliation?
```

<details><summary>Solutions</summary>

```{r message=FALSE, warning=FALSE}
# load packages
library(tidyverse)
library(stringr)
library(kableExtra)

# set seed for reproducibility
set.seed(1234)

# choose informative  prior distribution
# - I expect govts to last, on avg, 3 years, so fail at a 
#   rate of lambda = 1/3 per year, give-or-take about 1/10.
# - I know that the mean of the gamma is a/b and the sd is 
#   sqrt{a/(b^2)}, so I set a/b = 1/3 and sqrt{a/(b^2)} = 1/10
#   and solve. You could also just experiment with different 
#   values
alpha_inf <- 100/9
beta_inf <-  3*alpha_inf

# plot informative prior
ggplot() + xlim(0, 1) +
  stat_function(fun = dgamma, args = list(shape = alpha_inf, rate = beta_inf)) + 
  labs(title = "prior distribution", x = "lambda; expected failures/year", y = "prior density")

# choose **weakly informative** prior distribution
# - For this, I repeated the same process, but doubled my
#   give-or-take from 1/10 to 1/5,
alpha_weak <- 25/9
beta_weak <- 3*alpha_weak

# plot weakly informative prior
ggplot() + xlim(0, 1) +
  stat_function(fun = dgamma, args = list(shape = alpha_weak, rate = beta_weak)) + 
  labs(title = "prior distribution", x = "lambda; expected failures/year", y = "prior density")

# choose flattest prior distribution
# - I just want to make this prior as close to uniform as possible.
# - the failure rate (failures/year) is *surely* between 0 and 100
#   (lambda = 100 means that 100 governments are failing per year), 
#   so I plot the prior from zero to 100--it's basically flat   
alpha_flat <- 1
beta_flat <- 0.01

# plot flattish prior
ggplot() + xlim(0, 100) + ylim(0, NA) + 
  stat_function(fun = dgamma, args = list(shape = alpha_flat, rate = beta_flat)) + 
  labs(title = "prior distribution", x = "lambda; expected failures/year", y = "prior density")

# data
x <- c(142, 773, 333, 432, 1823)/365  # rescale to years

# make a data frame with the posterior and prior parameters
posts <- data.frame(prior_type = c("Informative", 
                                   "Weakly Informative",
                                   "Flat-ish"),
                    alpha_prior = c(alpha_inf, 
                                    alpha_weak,
                                    alpha_flat),
                    beta_prior = c(beta_inf, 
                                    beta_weak,
                                    beta_flat)) %>%
  # add posterior parameters
  mutate(alpha_posterior = alpha_prior + length(x),
         beta_posterior = beta_prior + sum(x)) %>%
  # create one row per parameter
  pivot_longer(cols = alpha_prior:beta_posterior) %>%
  # separate parameter from distribution type
  separate(name, into = c("parameter", "distribution_type")) %>%
  # put parameters into separate columns
  pivot_wider(names_from = parameter, values_from = value) %>%
  mutate(distribution_type = str_to_title(distribution_type),
         distribution_type = factor(distribution_type, levels = c("Prior", "Posterior"))) %>%
  mutate(prior_type = factor(prior_type, levels = c("Informative", "Weakly Informative", "Flat-ish")))

# compute point estimates and make a table
point_estimates <- posts %>% 
  #filter(distribution_type == "posterior") %>%
  group_by(prior_type, distribution_type) %>%
  mutate(mean = alpha/beta,
         median = median(rgamma(100000, alpha, beta)),
         mode = (alpha - 1)/beta,
         mean_expected_duration = mean(1/rgamma(100000, shape = alpha, rate = beta))) 

point_estimates %>%
  select(-alpha, -beta) %>%
  arrange(distribution_type, prior_type) %>%
  rename(`Prior` = prior_type, 
         `Distribution` = distribution_type,
         `Mean of lambda` = mean,
         `Median of lambda` = median,
         `Mode of lambda` = mode,
         `Mean of 1/lambda` = mean_expected_duration) %>%
  kable(format = "markdown", digits = 2)

# plot distributions
gg_posts <- posts %>%
  # add in lambda values for which to compute the prior and posterior density
  full_join(data_frame(lambda = seq(0, 2, length.out = 100)), 
            by = character()) %>%
  # compute the posterior and prior density for each lambda
  mutate(density = dgamma(lambda, shape = alpha, rate = beta))

ggplot(gg_posts, aes(x = lambda, y = density, 
                  color = prior_type, 
                  linetype = distribution_type)) + 
  geom_line()

```

</details>


## Method of Moments

### Mechanics

Suppose a random variable $X$. Then we refer to $E(X^k)$ as the $k$-th moment of the distribution or population. Similarly, we refer to $\text{avg}(x^k)$ as the $k$-th sample moment. 

For example, recall that $V(X) = E \left(X^2 \right) - \left[ E(X)\right]^2$. In example the variance of $X$ is the difference between the second moment and the square of the first moment. 

Recall that the law of large numbers guarantees that $\text{avg}(x) \xrightarrow[]{d} E(X)$. Thus, the first sample moment (the average) converges in probability to the first moment of $f$ (the expected value or mean). 

By the law of the unconscious statistician, we can similarly guarantee that $\text{avg}(x^k) \xrightarrow[]{d} E(X^k)$. Thus, the sample moments converge in distribution to moments of $f$. 

Now suppose that $f$ has parameters $\theta_1, \theta_2, ..., \theta_k$ so that $X \sim f(\theta_1, \theta_2, ..., \theta_k)$. We know (or can solve) for the moments of $f$ so that $E(X^1) = g_1(\theta_1, \theta_2, ..., \theta_k)$, $E(X^2) = g_2(\theta_1, \theta_2, ..., \theta_k)$, and so on. 

To use the method of moments, set the first $k$ sample moments equal to the first $k$ moments of $f$ and relabel $\theta_i$ as $\hat{\theta}_i$. Solve the system of equations for each $\hat{\theta}_i$.

### Example: Gamma Distribution

Suppose we have $N$ samples with replacement $x_n$ for $n \in \{1, 2, ..., N\}$ from a $\text{gamma}(\alpha, \beta)$ distribution. We can use the method of moments to derive estimators of $\alpha$ and $\beta$. We have two parameters, so we use the first two moments of $f$. For the gamma distribution, $E(X) = \frac{\alpha}{\beta}$ and $E(X^2) = V(X) + E(X)^2 = \frac{\alpha}{\beta^2} + \left( \frac{\alpha}{\beta} \right)^2 = \frac{\alpha(\alpha + 1)}{\beta^2}$. 

Since we have two parameters, we set the first two sample moments equal to the first two moments of the gamma distribution, add hats, and solve.

$$
\begin{aligned}
\displaystyle \frac{1}{N} \sum_{n = 1}^{N} x_n &=  \frac{\hat{\alpha}}{\hat{\beta}}\\
\displaystyle \frac{1}{N} \sum_{n = 1}^{N} x_n^2 &=  \frac{\hat{\alpha}(\hat{\alpha} + 1)}{\hat{\beta}^2}\\
\end{aligned}
$$

For simplicity, let $m_1 = \frac{1}{N} \sum_{n = 1}^{N} x_n$ and $m_2 = \frac{1}{N} \sum_{n = 1}^{N} x_n^2$. Then, solving for $\hat{\alpha}$ and $\hat{\beta}$, we obtain

$$
\begin{aligned}
\displaystyle \hat{\alpha} &=  \frac{m_1^2}{m_2 - m_1^2}\text{ and }\\
\displaystyle \hat{\beta} &=  \frac{m_1}{m_2 - m_1^2}\text{.}\\
\end{aligned}
$$

### Example: Toothpaste Cap Problem

We have a toothpaste cap--one with a wide bottom and a narrow top. We're going to toss the toothpaste cap. It can either end up lying on its side, its (wide) bottom, or its (narrow) top. 

We want to estimate the probability of the toothpaste cap landing on its top.

We can model each toss as a Bernoulli trial, thinking of each toss as a random variable $X$ where $X \sim \text{Bernoulli}(\pi)$. If the cap lands on its top, we think of the outcome as 1. If not, as 0. 

Suppose we toss the cap $N$ times and observe $k$ tops. How can we estimate the parameter $\pi$?

We can use the method of moments to estimate this single parameter. 

1. Set the first moment of the sample to the first moment of the Bernoulli distribution.
1. Add a hat to the quantities to estimate. 
1. Solve.

This process is nearly trivial for the Bernoulli distribution. $\text{sample average} = \frac{k}{N} = \hat{\pi}$. 

Thus, the fraction of the draws that land on their top is the methods of moments estimator of the probability of getting a top.

### Example: Population Average

Suppose we have $N$ samples with replacement $x_n$ for $n \in \{1, 2, ..., N\}$ from a population. We want to estimate the population average.

To estimate the population average, we would simply set the first sample moment $m_1 = \displaystyle \frac{1}{n} \sum_{n = 1}^{N} x_n^1$ equal to the first population moment $\mu_1 = E(X^1)$. Then we equate the sample moments to the population moments, add hats to the quantities of interest, and solve. Because $E(X^1)$ it the population average, there's not much work to do.

$$
\begin{aligned}
m_1 &= \mu_1\\
\displaystyle \frac{1}{N} \sum_{n = 1}^{N} x_n &=  \widehat{E(X)}\\
 &=  \text{estimate of population average}\\
\end{aligned}
$$

Once we solve for the quantity of interest, we can use that solution to *estimate* the quantity of interest. In this example, the sample average $\displaystyle \frac{1}{n} \sum_{n = 1}^{N} X_n$ is methods of moments estimator of the population average.

### Example: Population Average and SD

Suppose we also want the population SD. We set the second sample moment $m_2 = \displaystyle \frac{1}{n} \sum_{n = 1}^{N} x_n^2$ equal to the second population moment $\mu_2 = E(X^2)$. Then we equate the sample moments to the population moments and solve for the quantities of interest. 

$$
\begin{aligned}
\displaystyle \frac{1}{n} \sum_{n = 1}^{N} x_n &=  \widehat{E(X)}\\
\displaystyle \frac{1}{n} \sum_{n = 1}^{N} x_n^2 &=  \widehat{E(X^2)}\\
\end{aligned}
$$
Recall that the variance equals $V(X) = E \left( X^2 \right) - E \left( X \right)^2$ and the population $\text{SD}(X) = \sqrt{V(X)}$.

$$
\begin{aligned}
\displaystyle \sqrt{\frac{1}{N} \sum_{n = 1}^{N} x_n^2 - \left[ \displaystyle \frac{1}{N} \sum_{n = 1}^{N} x_n \right]^2} &=  \sqrt{\widehat{E(X^2)} - \widehat{E(X)}^2}\\
 &= \text{estimate of population SD}
\end{aligned}
$$
In this case, the *sample SD* is the methods of moments estimator for the population SD. 

*Notes*

- This is the same formula at the top of p. 74 of FPP. 
- Some authors recommend using $N - 1$ rather than $N$, so that $\sqrt{\frac{1}{N - 1} \sum_{n = 1}^{N} x_n^2 - \left[  \frac{1}{N - 1} \sum_{n = 1}^{N} x_n \right]^2}$. However, the formula using $N - 1$ is not the methods of moments estimator of the population SD. 


### Example: German Tank Problem

In World War II, the Germans manufactured some unknown number of tanks $\nu$, but gave each a sequential serial number $S = \{1, 2, ..., \nu\}$. It's of great strategic importance to know the $\nu$, the total number of tanks manufactured. 

When the Allies would destroy or capture a tank, they would record the serial number. Soon they had a data set of $N$ observed serial numbers $\{x_1, x_2, ..., x_N\}$, which they can treat as random   without replacement from $S$. 

Using the MOM, we can develop an estimator $\hat{\nu}_{MM}$ of the parameter $\nu$. We have only a single parameter $\nu$, so we match just the first sample moment $\frac{1}{N}\sum x^1_i = \text{avg}(x)$ with the first population moment $E(X) = \frac{\nu + 1}{2}$.

Then we have

$$
\begin{aligned}
\frac{1}{N} \sum_{n = 1}^{N} x_n &=  \frac{\hat{\nu}_{MM} + 1}{2}\\
\frac{2}{N} \sum_{n = 1}^{N} x_n &=  \hat{\nu}_{MM} + 1\\
\frac{2}{N} \left( \sum_{n = 1}^{N} x_n \right) - 1& =  \hat{\nu}_{MM}
\end{aligned}
$$

Here's a small simulation to demonstrate the estimator in repeated samples

```{r}
# the estimator
calc_nu_hat <- function(x) {
  N <- length(x)
  sum_x <- sum(x)
  nu_hat <- ((2 / N) * sum_x) - 1
  return(nu_hat)
}

# simulation parameters
n_sims <- 5  # the number of repeated samples
N <- 5       # the number of observations in each sample
nu <- 100    # the true value of nu

# do the simulation
set.seed(12345)
for (i in 1:n_sims) {
  # take the ith sample
  x <- sample(1:nu, size = N, replace = FALSE)
  x <- sort(x)  # sort x for convenience
  # compute the estimate
  nu_hat <- calc_nu_hat(x)
  # print the results
  data <- paste0(stringr::str_pad(x, 3, pad = "0"), collapse = ", ")
  estimate <- paste0("nu_hat = ", round(nu_hat, 2))
  cat("Sim. #", i, ": ", data, " --> ", estimate, "\n", sep = "")
}
```

Take a look at the 5 data sets above and the associated MOM estimates. Supposing you didn't know $\nu$, would these estimates seem reasonable? Can you spot any problems? (Look closely at Simulation #5.)

### Remarks

The method of moments has some advantages:

1. MOM intuitive; it just makes sense that the sample mean is a reasonable estimate of the population mean. 
1. MOM is numerically easy, usually. Because we're just computing (and sometimes then transforming) sample moments, we can usually compute these estimates quickly.
1. MOM demands few assumptions. We don't need to define the distribution of the data. It's sufficient, for example, to declare that it has a mean $mu$ and a variance $\sigma^2$ to obtain MOM estimator of these quantities.

However, the methods three disadvantages:

1. MOM estimators are rarely better and sometimes worse than maximum likelihood estimators.
1. MOM estimates sometimes fall outside the logically possible parameter space (see Simulation #5 of the German tank problem above) and sometimes outside the parameter space all together (e.g., negative estimates of parameters that can only be positive).

Because of the disadvantages, researchers rarely use the method of moments. However, they're a general, intuitive principle here: **(functions of) sample moments are reasonable estimates of (functions of) population moments**. In particular, 

1. $\frac{1}{n} \sum_{n = 1}^{N} x_n$ is a reasonable estimate of a population mean, and 
1. $\sqrt{\frac{1}{N} \sum_{n = 1}^{N} x_n^2 - \left[  \frac{1}{N} \sum_{n = 1}^{N} x_n \right]^2}$ is a reasonable estimate of the population SD.

### Exercises

```{exercise, label="mm-exponential"}
Suppose we collect $N$ random samples $x = \{x_1, x_2, ..., x_N\}$ and model each draw as a random variable $X \sim \text{expontial}(\lambda)$ with pdf $f(x_n | \lambda) = \lambda e^{-\lambda x_n}$. Find the method of moments estimator of $\lambda$. Hint: the mean of the exponential distribution is $\frac{1}{\lambda}.
```

<details><summary>Solution</summary>
Since we have a single parameter to estimate, we just set the first sample moment (the average) equal to the first moment of the distribution (the mean; $\frac{1}{\lambda}$) and add a hat. Thus, $\text{avg}(x) = \frac{1}{\hat{\lambda}}$ and therefore $\hat{\lambda} = \frac{1}{\text{avg}(x)}$.
</details>

```{exercise, label="mm-govt-duration"}
Model the data from Exercise \@ref(exr:govt-duration-bayes) as draws from an exponential distribution with rate parameter $\lambda$. Use the method of moments estimator from Exercise \@ref(exr:mm-exponential) to estimate $\lambda$. 

What if you want to estimate the mean duration $\mu = \frac{1}{\lambda}$ rather than the failure rate $\lambda$? Can you just use $\hat{\mu} = \frac{1}{\hat{\lambda}}$?
```

<details><summary>Solution</summary>
```{r}
x <- c(142, 773, 333, 432, 1823)/365  # rescale to years
1/mean(x)  # mm estimator of lambda
```

By construction, MOM estimators are invariant under transformation, so the MOM estimate of $g(\theta)$ can be found by transforming the MOM estimate of $\theta$, so $\widehat{g(\theta)} = g(\hat{\theta})$. For this problem, $\hat{\mu} = \frac{1}{\hat{\lambda}}$. 

</details>


## Maximum Likelihood

### Mechanics

Suppose we have a random sample from a distribution $f(x \mid \theta)$. We find the maximum likelihood (ML) estimator $\hat{\theta}$ of $\theta$ by maximizing the likelihood of the observed data with respect to $\theta$.

In short, we take the likelihood from Section \@ref(likelihood) and find the parameter $\theta$ that maximizes it. 

In practice, to make the math and/or computation a bit easier, we manipulate the likelihood function in three ways:

1. Relabel the likelihood function $f(x \mid \theta) = L(\theta)$, since it's weird to maximize with respect to a conditioning variable.
1. Work with $\log L(\theta)$ rather than $L(\theta)$. Because $\log()$ is a monotonically increasing function, the $\theta$ that maximizes $L(\theta)$ also maximizes $\log L(\theta)$. 

Suppose we have samples $x_1, x_2, ..., x_N$ from $f(x \mid \theta)$. Then the likelihood is $f(x \mid \theta) = \prod_{n = 1}^N f(x_n \mid \theta)$ and $\log L(\theta) = \sum_{n = 1}^N \log \left[ f(x_n \mid \theta) \right]$. The ML estimator $\hat{\theta}$ of $\theta$ is $\arg \max \log L(\theta)$.

In applied problems, we might be able to simplify $\log L$ substantially. Occasionally, we can find a nice analytical maximum. In many cases, we have a computer find the parameter that maximizes $\log L$.

### Example: Toothpaste Cap Problem

For the toothpaste cap problem, we have the following likelihood, which I'm borrowing directly from Section \@ref(likelihood).
$$
\text{the likelihood:  } f(x | \pi) = \pi^{k} (1 - \pi)^{(N - k)}, \text{where } k = \sum_{n = 1}^N x_n \\
$$
Then, we relabel. 

$$
\text{the likelihood:  } L(\pi) = \pi^{k} (1 - \pi)^{(N - k)}\\
$$
Then, we take the log and simplify. 

$$
\text{the likelihood:  } \log L(\pi) = k \log (\pi) + (N - k) \log(1 - \pi)\\
$$
To find the ML estimator, we find $\hat{\pi}$ that maximizes $\log L$. In this case, the analytical optimum is easy.

$$
\begin{aligned}
\frac{d \log L}{d\hat{\pi}} = y \left( \frac{1}{\hat{\pi}}\right) + (N - y) \left( \frac{1}{1 - \hat{\pi}}\right)(-1) &= 0\\
\frac{y}{\hat{\pi}} - \frac{N - y}{1 - \hat{\pi}} &= 0 \\
\frac{y}{\hat{\pi}} &= \frac{N - y}{1 - \hat{\pi}} \\
y(1 - \hat{\pi}) &= (N - y)\hat{\pi} \\
y - y\hat{\pi} &= N\hat{\pi} - y\hat{\pi} \\
y  &= N\hat{\pi} \\
\hat{\pi} &= \frac{y}{N} = \text{avg}(x)\\
\end{aligned}
$$
The ML estimator for the Bernoulli distribution is the fraction of successes, or, equivalently, the average of the Bernoulli trials. 

The collected data consist of 150 trials and 8 successes, so the ML estimate of $\pi$ is $\frac{8}{150} \approx 0.053$. Compare the ML estimate with the posterior mean, median, and mode above.

### Example: German Tank Problem

The German tank problem is an excellent example of an ML estimator because one can think through the logic intuitively rather than mathematically.

Recall that the Allies observe $N$ serial numbers of German tanks. By treating these serial numbers as samples without replacement from a set of sequential serial numbers $S = \{1, 2, ..., \nu\}$, they can estimate the total number of German tanks $\nu$.

Because we're sampling *without* replacement, the likelihood is quite complicated.

$$
f(x \mid \nu) = f(x_1) f(x_2 \mid x_1) ... f(x_N | x_{N - 1}, ..., x_2, x_1), \\
\text{where } x_i \neq x_j \text{ for } i \neq j.
$$

Remember that $\nu$ is the total number of tanks. So the chance that we observe $x_1$ first is $\frac{1}{\nu}$. The chance we observe $x_2$ second, given that we observe $x_1$ first is $\frac{1}{\nu - 1}$. The chance we observe $x_N$ $N$th is $\frac{1}{\nu - (N - 1)}$.

$$
L(\nu) = \frac{1}{\nu} \times \frac{1}{\nu - 1} \times ... \times \frac{1}{\nu - (N - 1)}
$$
It's clear, then, that in order to make $L$ as large as possible, we need to make $\nu$ as small as possible. However, notice that $\nu$ cannot be less than $\max \{x_1, ..., x_N\}$ (i.e., the largest serial number must be greater than or equal to the largest *observed* serial number).

So what's the smallest value of $\nu$ that satisfies $\nu \geq \max \{x_1, ..., x_N\}$? Of course it's $\hat{\nu} = \max \{x_1, ..., x_N\}$.

To illustrate, I simulate the same five data sets from above and compute the MM and ML estimates. Take a look at the five data sets and five estimates. Would you say that one estimator seems better? Can the ML estimate ever be too large? Can the MM estimate ever be too small?

```{r}
# the estimators
calc_nu_hat_mm <- function(x) {
  N <- length(x)
  sum_x <- sum(x)
  nu_hat <- ((2 / N) * sum_x) - 1
  return(nu_hat)
}
# the estimators
calc_nu_hat_ml <- function(x) {
  nu_hat <- max(x)
  return(nu_hat)
}

# simulation parameters
n_sims <- 5  # the number of repeated samples
N <- 5       # the number of observations in each sample
nu <- 100    # the true value of nu

# do the simulation
set.seed(12345)
for (i in 1:n_sims) {
  # take the ith sample
  x <- sample(1:nu, size = N, replace = FALSE)
  x <- sort(x)  # sort x for convenience
  # compute the estimates
  nu_hat_mm <- calc_nu_hat_mm(x)
  nu_hat_ml <- calc_nu_hat_ml(x)
  # print the results
  data <- paste0(stringr::str_pad(x, 3, pad = "0"), collapse = ", ")
  mm_estimate <- paste0("MM estimate = ", round(nu_hat_mm, 2))
    ml_estimate <- paste0("ML estimate = ", round(nu_hat_ml, 2))
  cat("Sim. #", i, ": ", data, " --> ", ml_estimate, "; ", mm_estimate, "\n", sep = "")
}
```

### Example: Poisson Distribution

Suppose we collect $N$ random samples $x = \{x_1, x_2, ..., x_N\}$ and model each draw as a random variable $X \sim \text{Poisson}(\lambda)$. Find the ML estimator of $\lambda$.

$$
\begin{aligned}
\text{Poisson likelihood: } f(x \mid \lambda) &= \prod_{n = 1}^N \frac{\lambda^{x_n} e^{-\lambda}}{x_n!} \\
L(\lambda) &= \prod_{n = 1}^N \frac{\lambda^{x_n} e^{-\lambda}}{x_n!} \\
\log L(\lambda) &= \sum_{n = 1}^N \log \left[ \frac{\lambda^{x_n} e^{-\lambda}}{x_n!} \right]\\
&= \sum_{n = 1}^N \left[ x_n \log \lambda + (-\lambda) \log e - \log x_n! \right]\\
&= \log \lambda \left[ \sum_{n = 1}^N x_n \right]  -N\lambda + \sum_{n = 1}^N \log (x_n!) \\
\end{aligned}
$$
To find the ML estimator, we find $\hat{\lambda}$ that maximizes $\log L$. In this case, the analytical optimum is easy.

$$
\begin{aligned}
\frac{d \log L}{d\hat{\lambda}} = \frac{1}{\hat{\lambda}} \left[ \sum_{n = 1}^N x_n \right] - N &= 0\\
\frac{1}{\hat{\lambda}} \left[ \sum_{n = 1}^N x_n \right] &= N \\
\left[ \sum_{n = 1}^N x_n \right] &= N \hat{\lambda} \\
\hat{\lambda} &= \frac{ \sum_{n = 1}^N x_n }{N} = \text{avg}(x)  \\
\end{aligned}
$$
The ML estimator for the Poisson distribution is just the average of the samples. 

### Remarks

The ML estimator is extremely common in political science because they are general, fast, and work extremely well. Lots of models that you've heard of, such as logistic regression, are estimated with ML.

We can even obtain ML estimates for the linear regression model. We assume that the observed data are samples from a normal distribution with mean $\mu_n = \alpha + \beta x_n$ and variance $\sigma^2$. For this model, the least-squares estimate that we learned earlier is also the ML estimates. 

### Exercises

```{exercise, label="ml-exponential"}
Suppose we collect $N$ random samples $x = \{x_1, x_2, ..., x_N\}$ and model each draw as a random variable $X \sim \text{expontial}(\lambda)$ with pdf $f(x_n | \lambda) = \lambda e^{-\lambda x_n}$. Find the maximum estimator of $\lambda$.
```

<details><summary>Solution</summary>
The math follows the Poisson example closely. However, the solution is the inverse--$\hat{\lambda} = \frac{N}{\sum_{n = 1}^N x_n } = \frac{1}{\text{avg}(x)}$.
</details>

```{exercise, label="ml-govt-duration"}
Model the data from Exercise \@ref(exr:govt-duration-bayes) as draws from an exponential distribution with rate parameter $\lambda$. Use the maximum estimator from Exercise \@ref(exr:ml-exponential) to estimate $\lambda$. 

What if you want to estimate the mean duration $\mu = \frac{1}{\lambda}$ rather than the failure rate $\lambda$? Can you just use $\hat{\mu} = \frac{1}{\hat{\lambda}}$?
```

<details><summary>Solution</summary>
```{r}
x <- c(142, 773, 333, 432, 1823)/365  # rescale to years
1/mean(x)  # ml estimator of lambda
```

Like MM estimators, ML estimators are invariant under transformation, though this is not as obvious as it is for MM estimators. For this problem, $\hat{\mu} = \frac{1}{\hat{\lambda}}$. 
</details>
